---
title: "Data HW1_Jiaqi Li"
author: "Jiaqi Li"
date: "April 7, 2019"
output: pdf_document
---

#Question 1
##1
```{r}
options(warn = -1)
#----------------part 1---------------------#
library(ggplot2)
data(diamonds)
qplot(carat,log(price),data = diamonds, col = cut, alpha = I(1/2),
      main = "log(price) vs. carat") + theme_bw() +
  scale_color_brewer(palette="Set3")
qplot(carat,sqrt(price),data = diamonds, col = cut, alpha = I(1/2),
      main = "log(price) vs. cut") + theme_bw() +
  scale_color_brewer(palette="Set3")
```

##2
```{r}
#----------------part 2---------------------#
library(dplyr)
require(gridExtra)
cutf = as.character(diamonds$cut) %>% as.factor()
diamonds$cut = cutf
out = lm(log(price)~carat+cut,data = diamonds)
qplot(out$fitted.values,out$residuals, alpha = I(1/2),
      col = diamonds$cut, main = "fitted vs. residuals")+
  theme_bw()
qplot(diamonds$carat,out$residuals, alpha = I(1/2),
      col = diamonds$cut, main = "carat vs. residuals")+
  theme_bw()
```

**Based on the observation, we can tell the residules plots show that at small fitted values, the linear regression line fits the data farily, but at the large fitted value, the linear regression line fits the data poorly. As the price increases, the residules also increase. This inplies that the linear regression fit the data poorly since the residules shows a trend.**


#Question 2
##1
```{r}
#----------------part 1---------------------#
library(data.table)
library(haven)
StockR = read_dta("StockRetAcct_insample.dta")
StockR = as.data.table(StockR)
StockR[,ExRet:=exp(lnAnnRet)-exp(lnRf)]
StockR[,lnIssue:=jitter(lnIssue,amount=0)]
for(i in 1981:2014){
  StockR[year == i, iss_vingtile_yr:=cut(StockR[year == i,]$lnIssue,
         breaks = quantile(StockR[year == i,]$lnIssue,probs = c(0:10)/10,na.rm = T),
         include.lowest=T,labels=F)]
}
Stock_R_Portf_yr = StockR[,list(MeanER_Yr = weighted.mean(ExRet,MEwt)),
                   by = list(iss_vingtile_yr,year)]
Stock_R_Portf = na.omit(Stock_R_Portf_yr[,list(MeanER = mean(MeanER_Yr)),
                by = iss_vingtile_yr])
```

##2
```{r}
#----------------part 2---------------------#
qplot(iss_vingtile_yr, MeanER, data = Stock_R_Portf,col=I("blue"),
      na.rm = T, main = "Average Return of Each Portfolio") + 
  geom_smooth(col=I("red"))+theme_bw()
```

**The pattern does not seems very linear, but it does show a slightly linear trend with negative slope.**


##3
```{r}
#----------------part 3---------------------#
StockR[iss_vingtile_yr == 1, transformed_iss:=-1]
StockR[iss_vingtile_yr == 10, transformed_iss:=1]
for(i in 2:9){
  StockR[iss_vingtile_yr == i, transformed_iss:=0]
}
out2 = lm(ExRet ~ transformed_iss, data = StockR)
summary(out2)
#this implies short stocks in decile 10 and long stocks in decile 1
```

**Based on the coefficient of the transformed issuance characteristic, we will earn an excess return if the characeristic is -1 and negative excess return is the characteristic is 1. Thus, we should long the portfolio contains stocks in Decile 1 and short the portfolio contains stocks in Decile 10.**


#Question 3
##1
```{r}
#----------------part 1---------------------#
StockR = read_dta("StockRetAcct_insample.dta")
StockR = as.data.table(StockR)
StockR[,ExRet:=exp(lnAnnRet)-exp(lnRf)]
StockR[,lnBM:=jitter(lnBM,amount=0)]
StockR[,lnME:=jitter(lnME,amount=0)]
for(i in 1981:2014){
  StockR[year == i, bm_quintile_yr := cut(StockR[year == i,]$lnBM,
         breaks = quantile(StockR[year == i,]$lnBM, probs = c(0:5)/5,
         na.rm = T), include.lowest = T, labels = F)]
  StockR[year == i, me_quintile_yr := cut(StockR[year == i,]$lnME,
         breaks = quantile(StockR[year == i,]$lnME, probs = c(0:5)/5,
         na.rm = T), include.lowest = T, labels = F)]
}
```

##2
```{r}
#----------------part 2---------------------#
Stock_R_Portf_double_yr = StockR[,list(MeanER_Yr = weighted.mean(ExRet,MEwt)),
  by = list(bm_quintile_yr,me_quintile_yr,year)]
Stock_R_Portf_double = na.omit(Stock_R_Portf_double_yr[,list(MeanER = mean(MeanER_Yr)),
  by = list(bm_quintile_yr,me_quintile_yr)])
setkey(Stock_R_Portf_double,me_quintile_yr)

for(i in 1:5){
  assign(paste("Plot",i,sep=""),
         qplot(bm_quintile_yr, MeanER,data = Stock_R_Portf_double[.(i)], na.rm = T,
         main = paste("Average Return of Portfolio with Size belongs to decile",i,sep=""),
         xlab="B/M Ratio", ylab = "Mean Excess Return") +
          geom_smooth(method = "lm",col=I("red"))+theme_bw())
}

p1= qplot(bm_quintile_yr, MeanER,data = Stock_R_Portf_double[.(1)], na.rm = T,
      main = "Mean Return of Size decile 1",
      xlab="B/M Ratio", ylab = "Mean Excess Return") +
  geom_smooth(method = "lm",col=I("red"))+theme_bw()
p2= qplot(bm_quintile_yr, MeanER,data = Stock_R_Portf_double[.(2)], na.rm = T,
      main = "Mean Return of Size decile 2",
      xlab="B/M Ratio", ylab = "Mean Excess Return") +
  geom_smooth(method = "lm",col=I("red"))+theme_bw()
p3= qplot(bm_quintile_yr, MeanER,data = Stock_R_Portf_double[.(3)], na.rm = T,
      main = "Mean Return of Size decile 3",
      xlab="B/M Ratio", ylab = "Mean Excess Return") +
  geom_smooth(method = "lm",col=I("red"))+theme_bw()
p4= qplot(bm_quintile_yr, MeanER,data = Stock_R_Portf_double[.(4)], na.rm = T,
      main = "Mean Return of Size decile 4",
      xlab="B/M Ratio", ylab = "Mean Excess Return") +
  geom_smooth(method = "lm",col=I("red"))+theme_bw()
p5= qplot(bm_quintile_yr, MeanER,data = Stock_R_Portf_double[.(5)], na.rm = T,
      main = "Mean Return of Size decile 5",
      xlab="B/M Ratio", ylab = "Mean Excess Return") +
  geom_smooth(method = "lm",col=I("red"))+theme_bw()
grid.arrange(p1,p2,p3,p4,p5)
```

**The assumption in linearity seems ok based on the 5 plots above. Only the portfolio of size in decile 4 and 5 shows a little bit non-linear trend, but overall, the average excess return does show a linear correlation with B/M ratio.**