---
title: "Data Analytics HW1_Jiaqi Li"
author: "Jiaqi Li"
date: "April 13, 2019"
output: pdf_document
---

#Problem Set 2

```{r}
options(warn = -1)
library(data.table)
library(ggplot2)
library(dplyr)
library(haven)
library(gridExtra)
library(lfe)
library(stargazer)
```

##Question 1
###1.
```{r}
#-------------------------------part 1-----------------------------#
data = read_dta("StockRetAcct_insample.dta") %>% as.data.table()
data[,ExRet:=exp(lnAnnRet)-exp(lnRf)]
setorder(data, FirmID, year)
time = unique(data$year)
p_ret = NULL
for(i in min(time):max(time)){
  temp = data[year == i,]
  temp_ret = lm(ExRet~lnInv, data = temp) %>% coef()
  p_ret = rbind(p_ret,temp_ret[2])
}
p_stat = list(mean_ret = mean(p_ret), std_ret = sd(p_ret),
              SR = mean(p_ret)/sd(p_ret),
              t_stat = sqrt(1+max(time)-min(time))*mean(p_ret)/sd(p_ret))
print(p_stat)
```

###2
$$ExRet_{i,t} = \delta_{0,t} + \delta_{1,t}lnInv_{i,t-1}$$
$\\$

$lnInv_{i,t-1} =$
$\begin{bmatrix}
1 & lnInv_{1,t-1} \\
. & . \\
. & . \\
. & . \\
1 & lnInv_{N,t-1}
\end{bmatrix}$
,
$ExRet_{t} =$
$\begin{bmatrix}
ExRet_{1,t-1} \\
. \\
. \\
. \\
ExRet_{N,t-1}
\end{bmatrix}$
,
then
$\begin{bmatrix}
\delta_{0,t} \\
\delta_{1,t}
\end{bmatrix}$
$= (lnInv_{t-1}'lnInv_{t-1})^{-1}lnInv_{t-1}'ExRet_t$

$\\$

$(lnInv_{t-1}'lnInv_{t-1})^{-1} = \frac{1}{N}\frac{1}{\frac{1}{N}\sum^{N}_{i=1}lnInv^2_{i,t-1}-(\frac{1}{N}\sum^N_{i=1}lnInv_{i,t-1})^2}$
$\begin{bmatrix}
\frac{1}{N}\sum^{N}_{i=1}lnInv^2_{i,t-1} & -\frac{1}{N}\sum^{N}_{i=1}lnInv_{i,t-1} \\
-\frac{1}{N}\sum^{N}_{i=1}lnInv_{i,t-1} & 1
\end{bmatrix}$

$\\$

$E_N[lnInv_{i,t-1}]=\frac{1}{N}\sum^{N}_{i=1}lnInv_{i,t-1}$,
$Var_N[lnInv_{i,t-1}]=\frac{1}{N}\sum^{N}_{i=1}lnInv^2_{i,t-1} - (\frac{1}{N}\sum^{N}_{i=1}lnInv_{i,t-1})^2$

$\\$

$(lnInv_{t-1}'lnInv_{t-1})^{-1} = \frac{1}{N}\frac{1}{Var_N[lnInv_{i,t-1}]}$
$\begin{bmatrix}
E_N[lnInv^2_{i,t-1}] & -E_N[lnInv_{i,t-1}] \\
-E_N[lnInv_{i,t-1}] & 1
\end{bmatrix}$

$\\$

$(lnInv_{t-1}'lnInv_{t-1})^{-1}lnInv_{t-1}'=\frac{1}{N}\frac{1}{Var_N[lnInv_{i,t-1}]}$
$\begin{bmatrix}
E_N[lnInv^2_{i,t-1}] & -E_N[lnInv_{i,t-1}] \\
-E_N[lnInv_{i,t-1}] & 1
\end{bmatrix}$
$\begin{bmatrix}
1 & . . . & 1\\
lnInv_{1,t-1} & . . . & lnInv_{N,t-1}
\end{bmatrix}$

$\\$

$\delta_{1,t} = \sum^{N}_{i=1}\frac{1}{N}\frac{lnInv_{i,t-1}-E_N[lnInv_{i,t-1}]}{Var_N[lnInv_{i,t-1}]}ExRet_{i,t}$

$\\$

$weight_{i,t-1} = \frac{1}{N}\frac{lnInv_{i,t-1}-E_N[lnInv_{i,t-1}]}{Var_N[lnInv_{i,t-1}]}$ ****************
$\\$(then 
$\delta_{1,t} = \sum^{N}_{i=1}weight_{i,t-1}ExRet_{i,t}$)
